Treatment heterogeneity in comparative effectiveness of teriparatide vs bisphosphonates: multi-database cohort study

### 2021-12-01 submission/manuscript.rmd
make negative control plot

\newpage
# Methods
## Study design

We performed a retrospective new user cohort study to estimate the effectiveness
of teriparatide compared to oral bisphosphonates in patients with
osteoporosis [@Ryan2013]. We included female participants above the age of 50
with established osteoporosis (any condition occurrence of hip, wrist, spine or
shoulder/humerus fracture in their history, prior to treatment initiation) and
at least 365 days of continuous observation period before the index event.
Patients were considered to be new-users if they received no anti-osteoporosis
drugs (raloxifene, bazedoxifene, denosumab, abaloparatide, romosozumab) 365 days
prior to treatment initiation with teriparatide or an oral bisphosphonate.

As a secondary analysis, we evaluated treatment effect heterogeneity using a risk-based approach. The

@@

- **IBM MarketScan Medicare Supplemental Database (MDCR)** is a claims database
  representing retirees in the United States with primary or Medicare
supplemental coverage.
- **Optum De-Identified Clinformatics® Data Mart Database – Date of Death
  (Optum-DOD)** is a claims database including members of private health
insurance.
- **Optum de-identified Electronic Health Record Dataset (Optum-EHR)** is an
  EHR database including Humedica’s Electronic Health Record.

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall treatment effect
estimates of teriparatide compared to oral bisphosphonates regarding the three
outcomes of interest. To account for potential measured confounding we
developed separate large-scale propensity score models within each database
based on LASSO logistic regression using the same predefined set of measured
covariates [@Tian2018]. We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate fitted in the 1:4
propensity score-matched subset of the considered patient population. Results
from different databases were summarized using random effects meta-analysis.

@@

\newpage
# Results

>>>>

\newpage
# Methods
## Study design
We performed a retrospective new user cohort study to estimate the
effectiveness of teriparatide compared to oral bisphosphonates in
patients with osteoporosis[8]. We included female participants above
the age of 50 with established osteoporosis (any condition occurrence
of hip, wrist, spine or shoulder/humerus fracture in their history,
prior to treatment initiation) and at least 365 days of continuous
observation period before the index event. Patients were considered to
be new-users if they received no anti-osteoporosis drugs (raloxifene,
bazedoxifene, denosumab, abaloparatide, romosozumab) 365 days prior to
treatment initiation with teriparatide or an oral bisphosphonate.

@@

- **IMB MarketScan Commercial Database (CCAE)** includes health
  insurance claims across the continuum of care (e.g. inpatient,
  outpatient, outpatient pharmacy, carve-out behavioral healthcare) as
  well as enrollment data from large employers and health plans across
  the United States who provide private healthcare coverage for more
  than 155 million employees, their spouses, and dependents. This
  administrative claims database includes a variety of fee-
  for-service, preferred provider organizations, and capitated health
  plans.
- **Optum De-Identified Clinformatics® Data Mart Database – Date of
  Death (Optum-DOD)**: is derived from a database of administrative
  health claims for members of large commercial and Medicare Advantage
  health plans. The database includes approximately 17-19 million
  annual covered lives, for a total of over 65 million unique lives
  over a 12 year period (1/2007 through 12/2019). The population is
  geographically diverse, spanning all 50 states.
- **Optum de-identified Electronic Health Record Dataset (Optum-EHR)**
  is derived from dozens of healthcare provider organizations in the
  United States, that include more than 700 Hospitals and 7000
  Clinics; treating more than 102 million patients receiving care in
  the United States.

##     Statistical analyses
We carried out two sets of analyses. First, we derived overall
treatment effect estimates of teriparatide compared to oral
bisphosphonates regarding the three outcomes of interest. To account
for potential measured confounding we developed separate large-scale
propensity score models within each database based on LASSO logistic
regression using the same predefined set of measured covariates
[@Tian2018]. We estimated treatment effects using Cox proportional
hazards models with treatment as the sole covariate fitted in the 1:4
propensity score-matched subset of the considered patient
population. Hazard ratios (HR) derived in different databases were
summarized using random effects meta-analysis.

@@

\newpage
# Results
## Treatment effect heterogeneity
